Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-23 ind_files/03_BD_processes.tex
Add some text discussing WNA moment eqns

\frac{d}{dt}\mathbb{E}[y_i^2] &= 2\mathbb{E}[y_iD_{i}] +  A_i^+(\boldsymbol{\alpha}(t)) = 2\mathrm{Cov}(y_i,D_i) + 2\mathbb{E}[y_i]\mathbb{E}[D_i]+A_i^+(\boldsymbol{\alpha}(t))\label{nD_moment_eqn_2nd_mom}

@@

Thus, whether stochastic fluctuations are expected to grow or decay is controlled by $D_i$, a measure of how the growth rate ($b_i - d_i$) changes along the direction of the fluctuation, whereas the spread of the fluctuations (the variance) has contributions from the net turnover rate ($A_i^+ = b_i + d_i$) and the covariance between the fluctuation and $D_i$. Note that unlike in the Price equation \eqref{nD_Price}, this is a true \emph{probability} covariance (as opposed to a statistical covariance between two deterministic quantities). In the case of the functional forms given by \eqref{nD_functional_forms_for_replicator}, we have:

@@

>>>>

\frac{d}{dt}\mathbb{E}[y_i^2] &= 2\mathbb{E}[y_iD_{i}] +  A_i^+(\boldsymbol{\alpha}(t)) = 2\langle y_i,D_i \rangle + 2\mathbb{E}[y_i]\mathbb{E}[D_i]+A_i^+(\boldsymbol{\alpha}(t))\label{nD_moment_eqn_2nd_mom}

@@

Where $\langle X, Y \rangle$ is the \emph{probability} covariance between two random variables $X$ and $Y$, defined as $\langle X, Y \rangle \coloneqq \mathbb{E}[XY] -  \mathbb{E}[X]\mathbb{E}[Y]$. This is not to be confused with the \emph{statistical} covariance defined by \eqref{nD_cov} that appears in the deterministic Price equation \eqref{nD_Price}. Thus, whether stochastic fluctuations are expected to grow or decay is controlled by $D_i$, a measure of how the growth rate ($b_i - d_i$) changes along the direction of the fluctuation, whereas the spread of the fluctuations (the variance) has contributions from the net turnover rate ($A_i^+ = b_i + d_i$) and the (probability) covariance between the fluctuation and $D_i$. In the case of the functional forms given by \eqref{nD_functional_forms_for_replicator}, we have:

@@

Thus, the expected behavior of fluctuations in the weak noise limit is controlled purely by fitness differences and mutational effects. If $\mathbb{E}[y_i] \equiv 0 \ \forall \ i$ is a stable fixed point for the system of equations defined by \eqref{nD_moment_eqn_mean_replicator}, then stochastic fluctuations are expected to decay, meaning that the deterministic point $\boldsymbol{\alpha}$ is a stable configuration for the complete dynamics when fluctuations are weak. In the case of 2-strategy matrix games (\emph{i.e.} when $m=2$ and the fitness functions $w_i(p)$ are of the form $w_i = \sum_j a_{ij}p_j$ for some constants $a_{ij}$), it has been shown that $\mathbb{E}[y_i] \equiv 0 \ \forall \ i$ is a stable fixed point for \eqref{nD_moment_eqn_mean_replicator} if and only if the population state $\boldsymbol{\alpha}$ is an evolutionarily stable strategy (ESS) for the determinstic game defined by \eqref{nD_det_limit} (which of course reduces to the replicator dynamics defined by \eqref{nD_replicator_mutator}), thus recovering a stochastic version of a classic result in evolutionary game theory~\citep{tao_stochastic_2007}.
